First of all I would like to express my honest gratitude to my advisor Tom
Fletcher for his guidance through my Ph.D. study. I was lucky to join Tom's group
in 2008. We decided to study the functional connectivity of the brain with fMRI
data, a relatively new research subject at that time, especially in Utah.  Tom's
patience and deep involvement is the key to why I had a smooth learning curve in
the early stage of the research work. More importantly, I learned from Tom the
methods of critical thinking and making scientific discoveries. These skills will
be my lifelong assets. I am particularly indebted to Tom's numerous reminders to
focus on the main theme of research, for I am often distracted by other
interesting but tangential research topics. From him I learned how to focus on
solving one problem in order to achieve a successful Ph.D.

I would like to give thanks also to Jeff Anderson, whose insight towards the
resting-state fMRI study sheds light on the later phase of my dissertation work. In
particular, Jeff proposed to validate the cross-session consistency with our
computational model, which becomes the last part of my dissertation and also a journal
paper currently under revision. Special thanks to Suyash Awate, for his many
ideas on the technical side during my second and third publications on the
Markov Chain Monte Carlo methods. I would also thanks Guido Gerig for his
intriguing comments during various seminars. His point of view is often at a
higher level, sometimes from the clinical side, which is especially helpful for a
computer science student. Also thanks to Tolga Tasdizen for his support during
my Ph.D process, and the thought-provoking questions during the qualifying exam and
proposal meeting.

I would also thank my fellow students and colleagues for their help whenever I
was stuck on a question and need someone to discuss with: Nikhil Singh,
Gopalkrishna Veni, Bo Wang, Xiang Hao, Jacob Hinkle, Neda Sadeghi, Sarang Joshi,
and numerous other people. This dissertation work has their direct or indirect
contribution. A big thanks to our SCI administrative staff Deborah Zemek, Edward
Cask, Magali Cohurn and Tony Portilo. Their support makes it possible for me to
focus on the research work, and makes the student labs at SCI such a fun place.
